---
title: 'STAT 632 : Homework 3'
author: "Pavan Malapati"
date: "2024-03-02"
output: pdf_document
---

Exercise 1. For this question use the Auto data set from the ISLR package. To access this data set first install the package using install.packages("ISLR") (this only needs to be done once). Then load the package into R using the command library(ISLR). You can read about this data set in the help menu by entering the command help(Auto).

(a) Make a scatter plot with mpg on the y-axis, and horsepower on the x-axis.

**Solution :**

```{r}
#install.packages("ISLR")
library(ISLR)
plot(Auto$horsepower,Auto$mpg,xlab = "Horsepower",ylab = "Miles per gallon")
```

(b) Use the lm() function to estimate a second degree (quadratic) polynomial regression model. That is, fit the model , where Y = mpg and x = horsepower. Use the summary() function to print the results.

**Solution :**

```{r}
fit <- lm(mpg ~ horsepower + I(horsepower^2), data = Auto)
summary(fit)
```

Y = 56.90 - 0.466x + 0.0012x^2 + e

(c) Use the fitted regression model to make a prediction and 95% prediction interval for the mpg of a vehicle that has horsepower = 150.

**Solution :**

```{r}
predict(fit,newdata=data.frame(horsepower = 150),data = Auto,interval = "prediction")
```

(d) Add the fitted second degree polynomial regression curve to the scatter plot of mpg versus horsepower. You may use either the base-R or ggplot2 approach.

**Solution :**

```{r}
library(ggplot2)
ggplot(data = Auto, aes(horsepower,mpg))+
geom_point()+
stat_smooth(method = 'lm', formula = y~poly(x,2), col = 'orange')
```

(e) Make a plot of the residuals versus fitted values, and a QQ plot of the standardized residuals. Comment on whether or not there are any violations of the assumptions for regression modeling.

**Solution :**

```{r}
plot(fit,which = 1,col='purple',lwd = 2)
```

```{r}
qqnorm(rstandard(fit))
qqline(rstandard(fit))
```

The variance is not constant and also the tails of residuals deviates from normal distribution

Exercise 2: For this question use the Carseats data set from the ISLR package.

(a) Fit a multiple linear regression model to predict Sales using Price, Urban, and US.

**Solution :**

```{r}
data(Carseats)
fit1 <- lm(Sales ~ Price + Urban + US, data = Carseats)
summary(fit1)
```

(b) Provide an interpretation of each coefficient in the model. Note that some of the variables are qualitative.

**Solution :**

-\> The intercept coefficient (beta0) represents the estimated mean value of Sales when Price, Urban, and US are all equal to zero.

-\> The Price coefficient (beta1) represents the estimated change in Sales for a one-unit increase in Price, holding Urban and US constant.

-\> The Urban coefficient (beta2) represents the estimated difference in Sales between stores located in urban areas versus stores located in rural areas, when Price and US are held constant.

-\> The US coefficient (beta3) represents the estimated difference in Sales between stores located in the US versus stores located in other countries, when Price and Urban are held constant.

(c) Write our the equation for the fitted model.

**Solution :** Sales = 13.043469 + (-0.054 Price) + (-0.021 UrbanYes) + (1.2 USYes)

(d) For which of the predictors can you reject the null hypothesis H0 : Beta j = 0 ?

**Solution :** From the summary output of the model, we can see that Price and USyes variables have p-values less than 0.05, indicating that we can reject the null hypothesis for these variables except **UrbanYes**. since its p value is not significant.

(e) On the basis of the your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.

**Solution :** We can fit a model that only includes Price and UsYes beacuse, these are the predictors for which there is evidence of association with Sales:

```{r}
fit2 <- lm(Sales ~ Price + US, data = Carseats)
summary(fit2)
```

(f) How well do the models in (a) and (e) fit the data?

**Solution :**

```{r}
# QQ plot for (A)
qqnorm(rstandard(fit1))
qqline(rstandard(fit1))

##QQ plot for (E)
qqnorm(rstandard(fit2))
qqline(rstandard(fit2))
```

The fits for both the models looks good and also residuals follows the normal distribution.

(g) Using the model from (e), obtain 95% confidence intervals for the coefficients.

**Solution :**

```{r}
confint(fit2,level = 0.95)
```
